Thread Management for High Performance Database Systems - Design and Implementation \\{\large Technical Report

### 2018-08-21 Presentations/Paper/paper_thread_pool.tex
Added pool_queue.pdf
Add pdf figures to paper tex file, and correct mail descriptions

\definecolor{mGreen}{rgb}{0,0.6,0}
\definecolor{mGray}{rgb}{0.5,0.5,0.5}
\definecolor{mPurple}{rgb}{0.58,0,0.82}
\definecolor{backgroundColour}{rgb}{1,1,1}

@@

\author{
	\IEEEauthorblockN{Robert Jendersie, Johannes Wuensche, Johann Wagner, \\Marten Wallewein-Eising, Marcus Pinnecke, Gunter Saake}
	\IEEEauthorblockA{\textit{Database and Software Engineering Group, Otto-von-Guericke University} \\
		Magdeburg, Germany \\
		\emph{firstname}.\emph{lastname}@ovgu.de} \and
}

@@

Parallel data processing can be achieved by different approaches, like instruction and data parallelism or multi threading. We focus on multi threading by implementing a thread pool for the graph database system \textsc{GeckoDB}\footnote{\textsc{GeckoDB} source code repository: \url{https://github.com/\textsc{GeckoDB}}}. The thread pool will be integrated into \textsc{Bolster}, a high performance library for parallel execution of primitives like for or filter on large data sets. \textsc{Bolster} has similarities to Intel \textsc{Tbb}~\cite{books/daglib/0018624}, a \textsc{C++} template library for parallel programming, but \textsc{Bolster} is written from scratch to perfectly fit into \textsc{GeckoDB}s specialized storage engine and vectorized query engine, both written in C11~\cite{pinneckegridtables}.

@@

\begin{figure*}[htbp]
	\centering \includegraphics[width=0.7\textwidth]{img/pool_structure.png}
	\caption{Design of the thread pool system}
	\label{fig0}
\end{figure*}

@@

\begin{figure}
	\centering\includegraphics[width=0.4\textwidth]{img/pool_queue.png}
	\caption{Workflow of task enqueueing and execution in the thread pool}
	\label{fig1}
\end{figure}

@@

\begin{figure}
\centering	\includegraphics[width=0.25\textwidth]{img/waitingconcept.png}
	\caption{Concept of waiting for tasks using a slot-map}
	\label{fig2}
\end{figure}

@@

\begin{figure}
	\includegraphics[width=0.5\textwidth]{img/pool_baseline.png}
	\caption{Time measures in nanoseconds of thread pool (blue) and baseline (orange) implementation with different numbers of threads.  }
	\label{fig6}
\end{figure}

@@

\begin{figure}
	\includegraphics[width=0.5\textwidth]{img/pool_avg.png}
	\caption{Relation between waiting time (orange) and completion time (blue) of thread pools with different sizes processing 10000 tasks}
	\label{fig7}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=0.4\textwidth]{img/slot_distr.png}
	\caption{Utilization \% with different slot distribution strategies}
	\label{fig8}
\end{figure}

\begin{figure}
	\centering\includegraphics[width=0.4\textwidth]{img/lock_free.png}
	\caption{Comparison of Utilization \% of the lock-free queue versus mutex based heap}
	\label{fig9}
\end{figure}

>>>>

\definecolor{mGreen}{rgb}{0,0.6,0}
\definecolor{mGray}{rgb}{0.5,0.5,0.5}
\definecolor{mPurple}{rgb}{0.58,0,0.82}
\definecolor{backgroundColour}{rgb}{1,1,1}
\definecolor{grayColor}{rgb}{0.56,0.56,0.56}

@@

\author{
	\IEEEauthorblockN{Robert Jendersie{\footnotesize\color{grayColor}$^\text{1}$}, Johannes Wuensche{\footnotesize\color{grayColor}$^\text{2}$}, Johann Wagner{\footnotesize\color{grayColor}$^\text{1}$}, \\Marten Wallewein-Eising{\footnotesize\color{grayColor}$^\text{2}$}, Marcus Pinnecke{\footnotesize\color{grayColor}$^\text{1}$}, and Gunter Saake{\footnotesize\color{grayColor}$^\text{1}$}}
	\IEEEauthorblockA{\textit{Database and Software Engineering Group, Otto-von-Guericke University} \\
		Magdeburg, Germany \\
		\emph{firstname}.\emph{lastname}@{\color{grayColor}(\footnotesize$^\text{1}$}ovgu {\color{grayColor}\footnotesize$\mid$ $^\text{2}$}st.ovgu{\color{grayColor})}.de}
}

@@

Parallel data processing can be achieved by different approaches, like instruction and data parallelism or multi threading. We focus on multi threading by implementing a thread pool for the graph database system \textsc{GeckoDB}\footnote{\textsc{GeckoDB} source code repository: \url{https://github.com/geckodb}}. The thread pool will be integrated into \textsc{Bolster}, a high performance library for parallel execution of primitives like for or filter on large data sets. \textsc{Bolster} has similarities to Intel \textsc{Tbb}~\cite{books/daglib/0018624}, a \textsc{C++} template library for parallel programming, but \textsc{Bolster} is written from scratch to perfectly fit into \textsc{GeckoDB}s specialized storage engine and vectorized query engine, both written in C11~\cite{pinneckegridtables}.

@@

\begin{figure*}[htbp]
	\centering \includegraphics[width=0.7\textwidth]{img/pool_structure.pdf}
	\caption{Design of the thread pool system}
	\label{fig0}
\end{figure*}

@@

\begin{figure}
	\centering\includegraphics[width=0.4\textwidth]{img/pool_queue.pdf}
	\caption{Workflow of task enqueueing and execution in the thread pool}
	\label{fig1}
\end{figure}

@@

\begin{figure}
\centering	\includegraphics[width=0.25\textwidth]{img/waitingconcept.pdf}
	\caption{Concept of waiting for tasks using a slot-map}
	\label{fig2}
\end{figure}

@@

\begin{figure}
	\includegraphics[width=0.5\textwidth]{img/pool_baseline.pdf}
	\caption{Time measures in nanoseconds of thread pool (blue) and baseline (orange) implementation with different numbers of threads.  }
	\label{fig6}
\end{figure}

@@

\begin{figure}
	\includegraphics[width=0.5\textwidth]{img/pool_avg.pdf}
	\caption{Relation between waiting time (orange) and completion time (blue) of thread pools with different sizes processing 10000 tasks}
	\label{fig7}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=0.4\textwidth]{img/slot_distr.pdf}
	\caption{Utilization \% with different slot distribution strategies}
	\label{fig8}
\end{figure}

\begin{figure}
	\centering\includegraphics[width=0.4\textwidth]{img/lock_free.pdf}
	\caption{Comparison of Utilization \% of the lock-free queue versus mutex based heap}
	\label{fig9}
\end{figure}
